---
title: "Exercise 2"
output: html_notebook
---

# Task 2

a)


```{r}

X_distributions = rbind(c(0.2,0.1,0.4,0.2,0.0,0.1), c(0.6,0.1,0.1,0.1,0.1,0.0), c(0.1,0.4,0.3,0.0,0.2,0.0))

set.seed(35)
draw_sample <- function(distributions) {
  rand_Y = sample(1:3, size=1, replace=TRUE, prob=c(0.4,0.3,0.3))
  distr = distributions[rand_Y,]
  X = expand.grid(0:1,0:2)[sample(1:6, 1, replace=TRUE, prob=distr),]
  c(rand_Y - 1, X[[1]], X[[2]])
}

test_set = matrix(unlist(lapply(1:100, function(x) draw_sample(X_distributions))), ncol = 3, byrow = TRUE)

Y0_instances = sum(test_set[,1] == 0) # Number of instances in class Y=0
X00_instances = sum((test_set[,2] == c(0,0)) + (test_set[,3] == 0) == 2) # Number of x=[0,0]
```

In class $Y=0$, there are `r Y0_instances` samples. In point $X = [0,0]$ there are `r X00_instances` samples

b)
Let's first count the number of instances for each class and each point. 

```{r}
X = matrix(c(0,0,0), c(0,0,0), nrow=3, ncol=2)
number_of_instances = list(X,X,X)
names(number_of_instances) <- c('Y0', 'Y1', 'Y2')

# Calculate number of instances for each class and point
apply(test_set, 1, function(x) {
  Y = x[1] + 1 # R starts indexes from 1
  X1 = x[2] + 1
  X2 = x[3] + 1
  
  number_of_instances[[Y]][X2,X1] <<- number_of_instances[[Y]][X2,X1] + 1
})
number_of_instances
```

Now, we can calculate smoothed estimates for each class using different smoothing parameters.

```{r}
smoothing <- function(x, smoothing_parameter = 1, total_samples_Y = 6) {
      (x + smoothing_parameter) / (total_samples_Y + 6*smoothing_parameter)
}

smoothed_estimates_ml = lapply(number_of_instances, function(Y) {
  total_samples_Y = sum(Y)
  matrix(sapply(Y, function(x) smoothing(x, smoothing_parameter = 0, total_samples_Y = total_samples_Y)), nrow=3, ncol=2)
})

smoothed_estimates_laplace = lapply(number_of_instances, function(Y) {
  total_samples_Y = sum(Y)
  matrix(sapply(Y, function(x) smoothing(x, total_samples_Y = total_samples_Y)), nrow=3, ncol=2)
})

smoothed_estimates_krichesky = lapply(number_of_instances, function(Y) {
  total_samples_Y = sum(Y)
  matrix(sapply(Y, function(x) smoothing(x, smoothing_parameter = 0.5, total_samples_Y = total_samples_Y)), nrow=3, ncol=2)
})
```

Smoothed estimates with ML-parameter:
```{r}
smoothed_estimates_ml
```

Smoothed estimates with Laplace-parameter:
```{r}
smoothed_estimates_laplace
```

Smoothed estimates with Krichesky-parameter:
```{r}
smoothed_estimates_krichesky
```